{\LARGE \textsc{Structural controlability of pollination networks}}

### 2016-05-04 paper/manuscript.Rnw
update partially results change code for randomisations

% title
% Article title
  \title{{\LARGE \textsc{Structural controlability of pollination networks}}}

@@

We started by quantifying the manageability of the analysed pollination communities. To do that, we calculated the maximum matching of the corresponding pollination network, and estimated the minimum number of species that need directed intervecntions to fully control the species abbundances in the community. If a network had more than one component (two species are in different components if there exist no path between them) we analysed the largest one. Although understanding the distribution of manageability across ecological networks is an useful result in itself, we wanted to test wheter perturbations, like invasion status, have an impact on manageability values. As, naively one might expect to see a negative relationship between diversity and manageability \autocite{Menge1995}, we used a set of generalised linear models to investigate the effect of invasion status while controling for \textit{(i)} the total number of species, \textit{(ii)} plant richness, \textit{(iii)} polinators richness, \textit{(iv)} the ratio of plants to pollinators richness, and \textit{(v)} the study site. Competing models were assessed by comparing the Akaike Information Criterion corrected for small sample sizes (AICc).

@@

\subsection*{Manageability}

@@

We found that if we were to control the dynamics of the whole community, we would need to control between \Sexpr{round(min(dplyr::filter(n_driver, type == "z-bi")$n_dr_p)*100)} and \Sexpr{round(max(dplyr::filter(n_driver, type == "z-bi")$n_dr_p)*100)}\% (mean \Sexpr{round(mean(dplyr::filter(n_driver, type == "z-bi")$n_dr_p)*100)}\%) of the species in the community if we were to harness bidirectional links between species, and between \Sexpr{round(min(dplyr::filter(n_driver, type == "weight")$n_dr_p)*100)} and \Sexpr{round(max(dplyr::filter(n_driver, type == "weight")$n_dr_p)*100)}\% (mean \Sexpr{round(mean(dplyr::filter(n_driver, type == "weight")$n_dr_p)*100)}\%) if we only take into account the strongest dependence. This corresponds to manageabilities between \Sexpr{round(min(1-dplyr::filter(n_driver, type == "z-bi")$n_dr_p), digits = 2)} to \Sexpr{round(max(1-dplyr::filter(n_driver, type == "z-bi")$n_dr_p), digits = 2)} and \Sexpr{round(min(1-dplyr::filter(n_driver, type == "weight")$n_dr_p), digits = 2)} to \Sexpr{round(max(1-dplyr::filter(n_driver, type == "weight")$n_dr_p), digits = 2)} respectively (Table S1). Despite different absolute values of manageability between these assuptions, the relative manageability of different communities was mantained (Spearman rank correlation \Sexpr{round(src_bi_w$estimate, digits = 2)}, $p = \Sexpr{as.numeric(format(src_bi_w$p.value, digits = 2))}$). In addition, the ratio of plant to pollinator species richness is positively correlated with manageability ($Pr(>\chi^2) = \Sexpr{as.numeric(format(anova(m8, m6, test = "LRT")$`Pr(>Chi)`[2], digits = 2))}$). Similarly, invaded communities tend to be more difficult to manage than uninvaded ones ($Pr(>\chi^2) = \Sexpr{as.numeric(format(anova(m8, m1, test = "LRT")$`Pr(>Chi)`[2], digits = 2))}$). In constrast, measures of biodiversity (like plant or pollinator richness) or the location of the study did not have a significant effect on the manageability (\autoref{fig:n_driver_explanatory}, Table S3).

\begin{figure}
\centering
<<fig explaining n_driver, echo = FALSE, warning = F, message = F, results = "asis", fig.height=2.2, fig.width=3.245>>=
p <- predict(m8, type = "response", se.fit = F)

n_driver_wb %>%
  dplyr::mutate(fit = predict(m8, type = "response", se.fit = T)$fit,
                se =  predict(m8, type = "response", se.fit = T)$se) %>%
  dplyr::filter(type == "z-bi" | type == "weight") %>% 
  ggplot(aes(x = n_pla/n_pol, y = 1-n_dr_p)) +
  geom_line(aes(y = 1-fit, colour = type, linetype = inv), size = 0.5) + 
  geom_point(aes(fill = type, shape = inv), show.legend = T, size = 1.25) +
  ylab("manegeability") + 
  xlab("plant / pollinator richness") +
  scale_fill_manual(values = c("#ffffff", "#bdbdbd"),
                      name = "direction of control", 
                      labels = c("asymmetry", "mutual dependence")) +
  scale_colour_manual(values = c("black", "#bdbdbd"),
                      name = "direction of control", 
                      labels = c("asymmetry", "mutual dependence")) + 
  scale_shape_manual(values = c(21, 23), 
                     name = "invasion status", 
                     labels = c("invaded", "non invaded")) +
  scale_linetype_manual(values = c(1, 12), 
                        name = "invasion status", 
                        labels = c("invaded", "non invaded")) +
  fer_theme +
  theme(plot.margin = grid::unit(c(5, 1, 2, 0), "mm"), 
        legend.position = "none", 
        # legend.title = element_blank(),
        legend.key = element_rect(fill = "white", 
                                  colour = "white"),
        legend.key.size = grid::unit(4, "mm"),
        legend.background = element_rect(colour = "black", size = 0.2),
        legend.box = "horizontal",
        legend.direction = "vertical") 
@
\caption{\label{fig:n_driver_explanatory} The relative manageability of a plant-pollinator community is largely influenced by the ratio of plant to pollinator species richness and to a smaller extent to the status of invasion. Invaded communities (circles, dashed lines) having larger levels of manageability. White and grey points correspond to manageabilities calculated using mutual dependences and interaction asymmetry respectively.}
\end{figure}

@@

\subsection*{Driver species}

@@

Species vary widely on the degree of their importance for network manageability. Our method to calculate the relative importance of species relies on calculating all possible maximum matchings. The number of matchings increasses exponentially with the number of links and nodes in a network. When we use mutual dependecies, there are two links between each interacting species, and we were only able to find the importance of different species in the eleven smallest networks. For the remaining nine networks, the number of maximum matchings is much larger than $4\times 10^{10}$ and the computation of the species relative importance became unfeasible.

Nevertheless, we were able to find the relative importance of species in the single link version of the networks where links were weighted by the assymetry. For the networks in which we were able to calculate the importance using mutual dependencies, we found a strong correlation between both approaches (Spearman's rank correlation ranged between $\rho = [\Sexpr{round(min(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V3, na.rm = T), digits = 2)}, \Sexpr{round(max(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V3, na.rm = T), digits = 2)}]$, median $\tilde{\rho} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V3, na.rm = T), digits = 2)})$. Indeed, when using the weighted rank correlation coefficient $r_{w2}$, we found that the agreement was stronger for highly ranked species than for lowly ranked species ($\tilde{r_{w2}} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V1, na.rm = T), digits = 2)}$ versus $\tilde{r_{w2}} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V2, na.rm = T), digits = 2)}$). Moreover both approaches identified a highly similar set of driver species (Jaccard similarity $J = [\Sexpr{round(min(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V4, na.rm = T), digits = 2)}, \Sexpr{round(max(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V4, na.rm = T), digits = 2)}]$, $\tilde{J} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V4, na.rm = T), digits = 2)})$

@@

Although plants showed significantly higher levels of importance for network control than pollinators (Mann-Withney-Wilcoxon test, $p \Sexpr{fix_small_pvalue(wilcox.test(pp ~ guild, data = dd, alternative = "greater")$p.value)}$), this difference can be largely explained by the strength of dependency of the species (the sum of the dependencies for a given species). Indeed, when averaging over the set of GLMMs that accounted at least 95\% of the evidencee based on AICc, we found that the strength of dependency  is the single most important factor determining the relative importance of species for network control (\autoref{table:driver_model}\todo{consider moving the table to thes supplementary}, \autoref{fig:driver_model_alt}). Contrastingly, the number of visits a species receives (or performs) and its contribution to nestedness have only a marginal effect on its relative importance for network control. In addition, in the ten communities that were invaded, invasive species were invariably classified as a driver species.

@@

The authors thank Dr. Takeuki Uno for the insight provided to find the set of all maximum matching algorithms, and Bernat Bramon, Matthew Hutchinson and Marilia Gaiarsa for feedback in early stages of the project. EFC acknowledges the support from the University of Canterbury Doctoral Scholarship, the University of Canterbury Meadow Mushroooms Postgraduate Scholarship, a travel grant from the European Space Agency and a Rutherford Discovery Fellowship (to DBS). DBS ackloledges the support of a Rutherford Discovery Scholarship, administered by the Royal Society of New Zealand.

>>>>

% title
% Article title
  \title{
  {\LARGE \textsc{Structural controlability of pollination networks}}\\
  {\LARGE \textsc{Structural controlability of ecological networks}}\\
  {\LARGE \textsc{Network manageability in the context of species invasion}}\\
  {\LARGE \textsc{Biotic invasions reduce manageability of ecological networks}}\\
  {\LARGE \textsc{Identifying the ``driver species'' of an ecosystem}}}

@@

We started by quantifying the manageability of the analysed pollination communities. To do that, we calculated the maximum matching of the corresponding pollination network, and estimated the minimum number of species that need directed intervecntions to fully control the species abbundances in the community. If a network had more than one component (two species are in different components if there exist no path between them) we analysed the largest one. Although understanding the distribution of manageability across ecological networks is an useful result in itself, we wanted to test wheter perturbations, like invasion status, have an impact on manageability values. As, naively one might expect to see a negative relationship between diversity and manageability \autocite{Menge1995}, we used a set of generalised linear models (with binomial error structure) to investigate the effect of invasion status while controling for \textit{(i)} the total number of species, \textit{(ii)} plant richness, \textit{(iii)} polinators richness, \textit{(iv)} the ratio of plants to pollinators richness, and \textit{(v)} the study site. Competing models were assessed by comparing the Akaike Information Criterion corrected for small sample sizes (AICc).

@@

@@

The studyed networks had levels of manageability between \Sexpr{round(min(1-dplyr::filter(n_driver, type == "z-bi")$n_dr_p), digits = 2)} and \Sexpr{round(max(1-dplyr::filter(n_driver, type == "z-bi")$n_dr_p), digits = 2)} (mean \Sexpr{round(mean(1-dplyr::filter(n_driver, type == "z-bi")$n_dr_p), digits = 2)}) when we considered bidirectional links weighted by the mutual dependences. This was in strong agreement  to the results obtained using the dependence asymmetry in which manageability ranged between \Sexpr{round(min(1-dplyr::filter(n_driver, type == "weight")$n_dr_p), digits = 2)} and \Sexpr{round(max(1-dplyr::filter(n_driver, type == "weight")$n_dr_p), digits = 2)} (Spearman rank correlation \Sexpr{round(src_bi_w$estimate, digits = 2)}, $p = \Sexpr{as.numeric(format(src_bi_w$p.value, digits = 2))}$). In addition we found that even when controling for diversity, invaded communities had were more difficult to manage than uninvaded ones. Out of the explored factors---total number of species, plant richness, pollinator richness and the ratio of plant to pollinator richness---only the ratio of plants to pollinator was retained in the final model (as determined by the AICc; \autoref{fig:n_driver_explanatory}, Table S3).

\begin{figure}
\centering
<<fig explaining n_driver, echo = FALSE, warning = F, message = F, results = "asis", fig.height=1.7, fig.width=3.245>>=

par_res <- as.data.frame(residuals(m8, "partial"))
names(par_res) <- paste0(make.names(names(par_res)), "_res")
par_res %<>%
  dplyr::bind_cols(m8$data) %>%
  dplyr::mutate(res_res = residuals(m8, "response"), 
                res_wor = residuals(m8, "working"))

par_res %>%
  ggplot(aes(x = inv, y = inv_res)) +
  # facet_wrap(~type) +
  geom_boxplot(size = 0.3,
               colour = "grey10", outlier.shape = 21, outlier.size = 1) +
  fer_theme +
  scale_x_discrete(labels = c("un-invaded", "invaded"), name = "") +
  ylab("adjusted manageability") +
  coord_flip() +
  theme(
        legend.position = "none",
				plot.margin = grid::unit(c(2, 1, 2, 0), "mm"), 
				# panel.grid.minor.x = element_blank(),
				panel.grid.minor.y = element_blank())

# p <- predict(m8, type = "response", se.fit = F)
# 
# n_driver_wb %>%
#   dplyr::mutate(fit = predict(m8, type = "response", se.fit = T)$fit,
#                 se =  predict(m8, type = "response", se.fit = T)$se) %>%
#   dplyr::filter(type == "z-bi" | type == "weight") %>% 
#   ggplot(aes(x = n_pla/n_pol, y = 1-n_dr_p)) +
#   geom_line(aes(y = 1-fit, colour = type, linetype = inv), size = 0.5) + 
#   geom_point(aes(fill = type, shape = inv), show.legend = T, size = 1.25) +
#   ylab("manegeability") + 
#   xlab("plant / pollinator richness") +
#   scale_fill_manual(values = c("#ffffff", "#bdbdbd"),
#                       name = "direction of control", 
#                       labels = c("asymmetry", "mutual dependence")) +
#   scale_colour_manual(values = c("black", "#bdbdbd"),
#                       name = "direction of control", 
#                       labels = c("asymmetry", "mutual dependence")) + 
#   scale_shape_manual(values = c(21, 23), 
#                      name = "invasion status", 
#                      labels = c("invaded", "non invaded")) +
#   scale_linetype_manual(values = c(1, 12), 
#                         name = "invasion status", 
#                         labels = c("invaded", "non invaded")) +
#   fer_theme +
#   theme(plot.margin = grid::unit(c(5, 1, 2, 0), "mm"), 
#         legend.position = "none", 
#         # legend.title = element_blank(),
#         legend.key = element_rect(fill = "white", 
#                                   colour = "white"),
#         legend.key.size = grid::unit(4, "mm"),
#         legend.background = element_rect(colour = "black", size = 0.2),
#         legend.box = "horizontal",
#         legend.direction = "vertical") 
@
\caption{\label{fig:n_driver_explanatory} Invaded communities have lower levels of manageability than uninvaded communities even when controling for the ratio of plant to pollinator richness. Adjusted manageability corrsponds to the partial working residuals of the invasion status.}
\end{figure}

@@

@@

We found a strong agreement between relative species' importance in the 11 networks for which we were able to calulate it using both mutual dependencies and asymmetries (Spearman's rank correlation ranged between $\rho = [\Sexpr{round(min(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V3, na.rm = T), digits = 2)}, \Sexpr{round(max(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V3, na.rm = T), digits = 2)}]$, median $\tilde{\rho} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V3, na.rm = T), digits = 2)})$. Indeed, when using the weighted rank correlation coefficient $r_{w2}$, we found that the agreement was stronger for highly ranked species than for lowly ranked species ($\tilde{r_{w2}} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V1, na.rm = T), digits = 2)}$ versus $\tilde{r_{w2}} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V2, na.rm = T), digits = 2)}$). Moreover both approaches identified a highly similar set of driver species (Jaccard similarity $J = [\Sexpr{round(min(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V4, na.rm = T), digits = 2)}, \Sexpr{round(max(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V4, na.rm = T), digits = 2)}]$, $\tilde{J} = \Sexpr{round(median(dplyr::filter(correlations, threshold == 0.5, type == "asymmetry", scaled == T)$V4, na.rm = T), digits = 2)})$. Therefore we employed the relative rankings obtained using asymmetries in the following analyses.

@@

Although plants showed significantly higher levels of importance than pollinators (Mann-Withney-Wilcoxon test, $p \Sexpr{fix_small_pvalue(wilcox.test(pp ~ guild, data = dd, alternative = "greater")$p.value)}$), this difference can be largely explained by the strength of dependency of the species (the sum of the dependencies for a given species). Indeed, when averaging over the set of GLMMs that accounted at least 95\% of the evidencee based on AICc, we found that the strength of dependency  is the single most important factor determining the relative importance of species for network control (\autoref{table:driver_model}\todo{consider moving the table to thes supplementary}, \autoref{fig:driver_model_alt}). In contrast, the number of visits a species makes (or receives) and its contribution to nestedness have only a marginal effect on its relative importance.

In addition, in the ten communities that were invaded, invasive species were invariably classified as a driver species.

@@

The authors thank Dr. Takeuki Uno for the insight provided to find the set of all maximum matching algorithms, and Bernat Bramon, Matthew Hutchinson and Marilia Gaiarsa for feedback in early stages of the project. EFC acknowledges the support from the University of Canterbury Doctoral Scholarship, the University of Canterbury Meadow Mushroooms Postgraduate Scholarship, a travel grant from the European Space Agency and a Rutherford Discovery Fellowship (to DBS). DBS ackloledges the support of a Rutherford Discovery Fellowship, administered by the Royal Society of New Zealand.
